Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms

### 2021-03-01 latex/MPP21/paper.tex
minor update to comments in latex
added text to section III, still need to add fig, performanc results, and write up analysis

\begin{enumerate}
	\item complexity of matmul discussion
	\item describe nature of approximation
	\item give explicit algorithm (Bini?)
	\item table of 12 algorithms
	\item accuracy plot?
\end{enumerate}

@@

\subsection{APA Algorithms}

@@

@@

\GB{reference PPoPP paper: \cite{BB15}}

\begin{enumerate}
	\item describe code generation and hybrid parallelization
	\item give performance plots (sequential and parallel)
\end{enumerate}

>>>>

@@

\subsection{APA Algorithms}
\label{sec:APAalgs}

@@

\GB{should we add an accuracy plot in this section?}

@@

We exploit the common algorithmic structure across APA (and exact) algorithms to develop a unified strategy of high performance multithreaded implementation using C++/OpenMP.
We build upon the work of Benson and Ballard \cite{BB15}, using code generation to apply the strategy to each algorithm.

\subsection{Hybrid Parallelization Strategy}

As described in \cref{sec:APAalgs}, each algorithm can be encoded by a set of triplets of coefficient matrices.
From this representation, we generate recursive code that computes the linear combinations for the inputs to each multiplication and combines the outputs.
The linear combinations are computed using a ``write-once'' strategy that was found to be most efficient in terms of memory bandwidth and performance.
The multiplications are either computed recursively or via a call to \texttt{gemm}, the interface to a highly efficient BLAS implementation of the classical algorithm.
Because we use only 1 recursive step in our experiments, there are no recursive calls, and every multiplication is performed by \texttt{gemm}.
We note that the original code generation tool was designed primarily for exact algorithms, and we generalized it to apply to all of the APA algorithms we consider here.

We adopt the ``hybrid'' parallelization strategy proposed by Benson and Ballard \cite{BB15}.
All linear combinations are parallelized in a straightforward way, in order to maximize the memory bandwidth of the machine.
The multiplications are parallelized as follows: given $r$ multiplications and $p$ threads, with $r=p\cdot q + \ell$ for integers $q$ and $\ell<r$, we assign each thread $q$ multiplications to be performed independently, and the remaining $\ell$ multiplications are performed by all threads using the multithreaded implementation of \texttt{gemm}.
\Cref{fig:hybrid} shows the hybrid strategy for $r=10$ (Bini's algorithm) and $p=4$.

\begin{figure}
\input{figs/hybrid}
\caption{Illustration of hybrid parallelization strategy for $r=10$ and 4 threads.}
\label{fig:hybrid}
\end{figure}

The hybrid strategy is efficient because each thread can achieve close to the peak performance of a core when computing independent matrix multiplications, even for relatively small problems.
The alternative strategy of employing multithreaded \texttt{gemm} for each of the $r$ multiplications (known as ``DFS'') suffers performance degradation for small problems, where the parallel implementation attains a smaller fraction of peak.
The hybrid strategy also perfectly load balances the computation across threads, as opposed to the alternate strategy of assigning the $\ell$ remainder multiplications to $\ell$ different threads (known as ``BFS''), leaving the other $p-\ell$ threads idle.

\subsection{Sequential Performance}

\GB{add sequential performance results for all 12 algorithms, compared to \texttt{gemm}}

\subsection{Parallel Performance}

\GB{add parallel performance results for all 12 algorithms, compared to \texttt{gemm}}
